Global well-being and mental health in the internet age

### 2022-10-26 ms.Rmd
Major revision

---
title: Global well-being and mental health in the internet age
shorttitle: Global well-being
leftheader: Global well-being
author: 
  - name: Matti Vuorre
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
    role:
      - Conceptualization
      - Data curation
      - Formal analysis
      - Investigation
      - Methodology
      - Project Administration
      - Software
      - Validation
      - Visualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name: Andrew K. Przybylski
    affiliation: 1
    corresponding: yes
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
    role:
      - Conceptualization
      - Data curation
      - Funding acquisition
      - Investigation
      - Methodology
      - Project Administration
      - Resources
      - Supervision
      - Visualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing

affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford

authornote: |
  \noindent \textbf{This pre-print is not yet peer-reviewed.}

  This research was supported by the Huo Family Foundation and Economic and Social Research Council Grant ES/T008709/1. All code supporting this manuscript is available at <https://doi.org/10.5281/zenodo.7004053>. This draft is available at <https://doi.org/10.31234/osf.io/9tbjy>.

abstract: |

  \textbf{Abstract.} In the last two decades the rise of internet technologies have impacted nearly all domains of human life and inspired concern that their widespread adoption has had a harmful effect on mental health and psychological well-being. However, research on the topic is contested and hampered by methodological shortcomings leaving the broader consequences of internet adoption unknown. We show that the past two decades have seen only small and inconsistent changes in global well-being and mental health. We demonstrate that the broadest available health and internet technology data does not support the idea that the adoption of internet and mobile broadband has been psychologically harmful on a global level. Moreover, little or no negative associations were observed for adolescents—a group thought particularly vulnerable. Further investigation of this topic requires transparent study of online behaviours where they occur: on online communication, media, social media, and video game platforms. We place an urgent call to increase the collaborative efforts between independent scientists and the internet technology sector.

  \textbf{Statement of Relevance} tbd

keywords: well-being, mental health, internet technology, technology effects
# wordcount: 
bibliography: "references.bib"
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
# header-includes:

csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(readxl)
library(imputeTS)
library(scales)
library(janitor)
library(haven)
library(labelled)
library(countrycode)
library(plotly)
library(lubridate)
library(kableExtra)
library(brms)
library(knitr)
library(here)
library(bayestestR)
library(ggh4x)
library(tidybayes)
library(ggdist)
library(DT)
library(tidyverse)

opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  include = knitr::is_html_output(),  # Show code only in HTML
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)
# For tables: Show NAs as blanks
options(knitr.kable.NA = '')

# Create directory for intermediate files
dir.create(here("data"), FALSE)

# Plotting options
theme_set(
  theme_linedraw(
    base_size = 10
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.6)),
      axis.ticks = element_line(size = rel(0.3))
    )
)
```

```{r functions}
#' Custom kableExtra formatting
#'
#' @param a table from kbl()
#'
#' @return a table
kable_custom <- function(x) {
  kable_classic(
    kable_input = x,
    # html_font = "Arial",
    font_size = 11,
    position = "center",
    full_width = FALSE
  )
}

# Function to replace extreme percentages
percent2 <- function(x, accuracy = .1) {
  x <- percent(x, accuracy = accuracy)
  x <- if_else(x == "100.0%", ">99.9%", x)
  x <- if_else(x == "0.0%", "<0.1%", x)
  x
}

# Function to align positive and negative values
number2 <- function(x, accuracy = .001, pad = FALSE) {
  out <- number(
    x,
    accuracy = accuracy,
    style_negative = "hyphen"
    )

  # Justify on decimal point
  if (pad) {
    # Add white space to start of non-negative numbers
    out[str_starts(out, "[0-9]")] <- str_c(
      " ", 
      out[str_starts(out, "[0-9]")]
    )
  }
  out

}
```

@@

# First liberate the data from the slow SPSS file and save to disk
gwp_path <- here("data-raw/Gallup/gwp-processed.rds")
if (!file.exists(gwp_path)) {
  gwp <- read_spss(
    here("data-raw/Gallup/The_Gallup_101521.sav"),
    col_select = c(
      YEAR_CALENDAR,
      COUNTRYNEW,
      WPID, WP1220, WP1219,
      WP16,
      WP60, WP61, WP63, WP65, WP67,
      WP68, WP69, WP70, WP71, WP74
    )
  )

  # Get rid of SPSS attributes
  gwp <- gwp %>%
    zap_labels() %>%
    zap_label() %>%
    zap_widths() %>%
    zap_formats()

  # Save into a good format
  write_rds(gwp, gwp_path)
} else {
  gwp <- read_rds(gwp_path)
}

@@

# Scale scores note rescaling
gwp <- gwp %>%
  mutate(
    Life_satisfaction = Life_satisfaction / 10,
    Negative_experiences =
      rowMeans(select(., wp68:wp74), na.rm = TRUE) - 1,
    Positive_experiences =
      rowMeans(select(., wp60:wp67), na.rm = TRUE) - 1
  ) %>%
  select(-c(wp60:wp74))

@@

# Check what names were changed and if any don't have harmonised counterpart
gwp %>%
  distinct(country) %>%
  arrange(country) %>%
  mutate(
    country_harmonised = countryname(country, destination = "cldr.short.en")
  ) %>%
  filter(country != country_harmonised | is.na(country_harmonised)) %>%
  kbl() %>%
  kable_custom()

@@

gwp_mock_path <- here("data-raw/gwp-MOCK.rds")
if (dir.exists(here("data-raw/Gallup")) & !file.exists(gwp_mock_path)) {
  library(synthpop)
  gwp_syn <- gwp %>% 
    drop_na()
  gwp_syn <- syn(
    gwp_syn, 
    method = c("", "", "", "", "", "", "normrank", "normrank"),
    maxfaclevels = 168
  )
  gwp_syn <- tibble(gwp_syn$syn)
  saveRDS(gwp_syn, gwp_mock_path)
}

# Load if actual data doesn't exist
if (!dir.exists(here("data-raw/Gallup"))) {
  cat("Using synthetic GWP data.")
  gwp <- read_rds(gwp_mock_path)
} else {
  cat("Using actual GWP data.")
}

# Load data files and merge to one table
gbd <- list.files(
  here("data-raw/GBD/"),
  pattern = ".zip", full.names = TRUE, recursive = TRUE
) %>%
  read_csv()

# First step is to clean names
gbd <- clean_names(gbd)

# Confirm that correct measures were downloaded:
# distinct(gbd, measure, metric)
gbd <- select(gbd, -metric, -measure)
gbd <- gbd %>%
  mutate(across(c(val, upper, lower), ~ .x / 100000))

@@

# Check what names were changed and if any don't have harmonised counterpart
gbd %>%
  distinct(country) %>%
  arrange(country) %>%
  mutate(
    country_harmonised = countryname(country, destination = "cldr.short.en")
  ) %>%
  filter(country != country_harmonised | is.na(country_harmonised)) %>%
  kbl() %>%
  kable_custom()
# Harmonise old names and replace only if harmonised name found
gbd <- gbd %>%
  mutate(
    country_harmonised = countryname(
      country,
      destination = "cldr.short.en"
    )
  ) %>%
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>%
  select(-country_harmonised)

# Download data sets from the ITU website if not yet downloaded
if (!file.exists(here("data-raw/ITU/PercentIndividualsUsingInternet.xlsx"))) {
  dir.create("data-raw/ITU")
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/December/PercentIndividualsUsingInternet.xlsx",
    here("data-raw/ITU/PercentIndividualsUsingInternet.xlsx")
  )
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/December/FixedBroadbandSubscriptions_2000-2020.xlsx",
    here("data-raw/ITU/FixedBroadbandSubscriptions_2000-2020.xlsx")
  )
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/December/MobileBroadbandSubscriptions_2007-2020.xlsx",
    here("data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx")
  )
}

# Read and reshape fixed broadband data
itu_fixed <- read_xlsx(
  here("data-raw/ITU/FixedBroadbandSubscriptions_2000-2020.xlsx"),
  sheet = 2,
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Country)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(fixed = as.numeric(value)) %>%
  select(country = Country, year, fixed)

# Read and reshape mobile broadband data
itu_mobile <- read_xlsx(
  here("data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx"),
  sheet = 2,
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Country)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(mobile = as.numeric(value)) %>%
  select(country = Country, year, mobile)

# Percent using internet
itu_percent <- read_xlsx(
  here("data-raw/ITU/PercentIndividualsUsingInternet.xlsx"),
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Country)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(internet = as.numeric(value)) %>%
  select(country = Country, year, internet)

# Merge to one table in wide format (oldest on left to include all years)
itu <- reduce(list(itu_percent, itu_fixed, itu_mobile), left_join)

# values (1-100) to proportions (0-1) and years to numbers
itu <- itu %>%
  mutate(across(c(internet, fixed, mobile), function(x) x / 100)) %>%
  mutate(year = as.numeric(year))

# Drop tables
rm(itu_fixed, itu_mobile, itu_percent)

# Harmonise countries
# Check what names were changed and if any don't have harmonised counterpart
itu %>%
  distinct(country) %>%
  arrange(country) %>%
  mutate(
    country_harmonised = countryname(country, destination = "cldr.short.en")
  ) %>%
  filter(country != country_harmonised | is.na(country_harmonised)) %>%
  kbl() %>%
  kable_custom()

@@

# Summary
itu_long %>%
  group_by(name) %>%
  mutate(min_year = if_else(name == "mobile", 2007, 2000)) %>%
  summarise(
    n_miss_orig = sum(is.na(value) & year >= min_year),
    p_miss_orig = percent(n_miss_orig / sum(year >= min_year), .1),
    n_miss_imp = sum(is.na(value_i) & year >= min_year),
    p_miss_imp = percent(n_miss_imp / sum(year >= min_year), .1),
    imputed = n_miss_orig - n_miss_imp
  ) %>%
  kbl() %>%
  kable_custom()

@@

regions %>% 
  group_by(region) %>%
  summarise(
    N = n(),
    Countries = str_c(country, collapse = ", ")
  ) %>% 
  kbl() %>% 
  kable_custom()

# Create lagged predictors
itu <- itu %>% 
  arrange(region, country, year) %>%
  group_by(region, country) %>%
  mutate(
    across(
      c(internet, mobile),
      list(`1` = ~ lag(., 1)),
      .names = "{str_sub(.col, 1, 1)}{.fn}"
    )
  ) %>% 
  ungroup()

# Centering predictors within and between countries
itu <- itu %>%
  # Grand mean centering so both components are zero-centered
  mutate(across(c(i1, m1), ~. - mean(., na.rm = TRUE))) %>% 
  group_by(country) %>%
  mutate(
    # Between
    i1_cb = mean(i1, na.rm = TRUE),
    m1_cb = mean(m1, na.rm = TRUE),
    # Within
    i1_cw = i1 - i1_cb, 
    m1_cw = m1 - m1_cb
  ) %>% 
  ungroup()

@@

# Center and scale year
dat <- dat %>%
  mutate(year = (year - 2010) / 10)

# Identify values to use in MV models
# Indicate unique ITU values for MV model
dat <- dat %>%
  # Unique values exist for each country-year, but must be given for each
  # outcome
  group_by(region, country, year, outcome) %>%
  mutate(itu = 1:n() == 1, itum = 1:n() == 1) %>%
  ungroup()

# This prevents a glitch where rows where ITU==TRUE but internet is missing would not be included in model of val
dat <- dat %>% 
  mutate(
    itu = if_else(itu & is.na(internet), FALSE, itu),
    itum = if_else(itum & is.na(mobile), FALSE, itum)
    )

@@

# Model outcomes on the percentage (0-100) scale. Note internet is still 0-1
dat <- dat %>%
  mutate(across(c(val, se), ~ . * 100))

@@

# Selfharm models don't converge with SEs
dat <- dat %>% 
  mutate(se = if_else(outcome == "Selfharm", 0, se))
```

```{r data-describe}
bind_rows(
  "GWP" = gwp,
  "GBD" = gbd,
  .id = "Dataset"
) %>% 
  mutate(Dataset = fct_inorder(Dataset)) %>% 
  group_by(Dataset, region, outcome, country) %>% 
  summarise(
    n_n = sum(n),
    n_year = length(unique(year))
  ) %>% 
  summarise(
    N = sum(n_n), 
    Years = mean(n_year)
  ) %>%
  summarise(
    N_sum = round(mean(N)), 
    Years_mean = round(mean(Years))
  ) %>%
  kbl() %>% 
  kable_custom()
tmp <- gwp %>% 
  group_by(outcome) %>% 
  summarise(n=sum(n)) %>% 
  summarise(N=comma(median(n))) %>% 
  pull(N)

bind_rows(
  "GWP" = gwp,
  "GBD" = gbd,
  .id = "Dataset"
) %>% 
  mutate(Dataset = fct_inorder(Dataset)) %>% 
  count(Dataset, outcome) %>% 
  bind_rows(
    itu %>% 
      select(internet, mobile) %>% 
      pivot_longer(everything(), names_to = "outcome") %>% 
      drop_na(value) %>% 
      count(outcome) %>% 
      mutate(Dataset = "ITU")
  ) %>% 
  kbl() %>% 
  kable_custom()

gwp %>% 
  # filter(age %in% c("15 to 19", "50 to 54", "85 to 89")) %>% 
  group_by(age) %>% 
  summarise(
    n = quantile(n, probs = c(.1, .5, .9)),
    quantile = c("10%", "50%", "90%")
  ) %>% 
  pivot_wider(names_from = quantile, values_from = n) %>% 
  kbl() %>% 
  kable_custom()

itu %>%
  select(region, country, year, internet) %>%
  mutate(internet = internet * 100) %>%
  pivot_wider(names_from = year, values_from = internet) %>%
  select(-country) %>%
  tbl_summary(
    by = region,
    missing = "no",
    statistic = list(all_continuous() ~ "{mean}% ({N_nonmiss})")
  ) %>%
  add_overall(col_label = "Total") %>%
  as_kable_extra() %>%
  row_spec(0, bold = TRUE) %>%
  kable_custom()

if (!file.exists(here("data-raw/ITU/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx"))) {
  download.file("https://www.itu.int/en/ITU-D/Statistics/Documents/facts/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx", here("data-raw/ITU/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx"))
}
# Individuals using the internet per 100 inhabitants | World
tmp <- read_excel("data-raw/ITU/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx", range = "T85:AI86")
tmp %>% 
  kbl() %>% 
  kable_custom()
```

```{r models-process}
# Create a table of outcomes in appropriate order, so can map functions over this
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
fits <- tibble(outcome = oo) %>% 
  mutate(outcome = fct_inorder(outcome))

# The files are huge and unwieldy to work with. We create smaller files for each by dropping samples of the age:country interaction random effects. It is then easier to post-process the models.
dir.create("models-small", FALSE)

replace_chain <- function(x) {
  x %>%
    select(
      -starts_with("r_1"),
      -starts_with("r_2"),
      -starts_with("r_3"),
      -starts_with("z_1"),
      -starts_with("z_2"),
      -starts_with("z_3"),
      -starts_with("r_age:country"),
      -starts_with("L_")
    )
}

replace_chain_all <- function(x) {
  if (!file.exists(str_glue("models-small/{x}"))) {
    message(x)
    out <- readRDS(str_glue("models/{x}"))
    out$fit@sim$samples <- map(1:4, ~replace_chain(out$fit@sim$samples[[.]]))
    saveRDS(out, str_glue("models-small/{x}"), compress = FALSE)
  }
}

walk(
  list.files("models"),
  ~replace_chain_all(.x)
)

parameter_summary <- function(outcome, model) {
  path <- str_glue("models-small/brm-{outcome}-{model}.rds")
  as_draws_df(
    readRDS(path), 
    variable = c("^b_", "^sd_", "^cor_"), 
    regex = TRUE
  ) %>%
    summarise_draws(
      mean, sd,
      samples = length, 
      default_convergence_measures()
    )
}
parameter_summary <- memoise(parameter_summary, cache = cd)

convergence_summary <- tibble(
  path = list.files("models-small/", pattern = ".rds$", full.names = FALSE)
) %>% 
  separate(path, c("brm", "outcome", "model"), sep = "-") %>% 
  select(-brm) %>% 
  mutate(outcome = factor(outcome, levels = oo)) %>% 
  arrange(outcome) %>% 
  mutate(model = str_extract(model, "[0-9]")) %>% 
  mutate(parameters = map2(outcome, model, ~parameter_summary(.x, .y))) %>% 
  unnest(parameters)

convergence_summary %>% 
  filter(abs(rhat) > 1.03) %>% 
  arrange(outcome, model, variable) %>% 
  kbl(digits = c(rep(2, 7), 0, 0)) %>% 
  kable_custom()

# Function to draw a pairsplot of variables `v` of model in `path` 
dir.create(here("figures/diagnostics"), FALSE)
pairsplot <- function(filename, v) {
  out_path <- here(str_glue("figures/diagnostics/{basename(filename)}.png"))
  if (!file.exists(out_path)) {
    x <- readRDS(here(filename))
    agg_png(
      out_path,
      width = W*1.2, height = W*1.2, units = "in", res = 160
    )
    pairs(x$fit, pars = v)
    title(sub = out_path)
    dev.off()
  }
}

# Draw pairsplot for 10 worst rhats per model
convergence_summary %>% 
  group_by(outcome, model) %>% 
  arrange(outcome, model, desc(rhat)) %>% 
  slice(1:10) %>% 
  ungroup() %>% 
  group_by(outcome, model) %>% 
  summarise(v = list(variable)) %>% 
  mutate(path = str_glue("models-small/brm-{outcome}-{model}.rds")) %>% 
  ungroup() %>% 
  {walk2(.$path, .$v, ~pairsplot(.x, .y))}
```

```{r results}
# Before drawing any plots, we need to extract/compute the relevant parameters from the models. We do that here: For each outcome, we get the countries' average changes over time (in the outcome, internet, and mobile) and associations with internet and mobile
country_post_fun <- function(outcome) {

>>>>

---
title: Global well-being and mental health in the internet age
shorttitle: Global well-being
leftheader: Global well-being
author: 
  - name: Matti Vuorre
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
  - name: Andrew K. Przybylski
    affiliation: 1
    corresponding: yes
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  \noindent \textbf{This pre-print is not yet peer-reviewed.}
abstract: | 
  In the last two decades the rise of internet technologies have impacted nearly all domains of human life and inspired concern that their widespread adoption has impacted mental health and psychological well-being. However, research on the topic is contested and hampered by methodological shortcomings leaving the broader consequences of internet adoption unknown. We show that the past two decades have seen small and inconsistent changes in global well-being and mental health. The broadest available health and internet technology data does not support the idea that the adoption of internet and mobile broadband is consistently linked to negative psychological outcomes. However, associations between internet technology adoption and depression and self-harm were strongest for individuals aged between 10 and 34. Further investigation of this topic requires transparent study of online behaviours where they occur, namely on online platforms. We call for increased collaborative efforts between independent scientists and the internet technology sector.

keywords: well-being, mental health, internet technology, technology effects
wordcount: '`r scales::number(wordcountaddin:::word_count("ms.Rmd"), big.mark = ",")`'
bibliography: "references.bib"
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
# header-includes:
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(readxl)
library(imputeTS)
library(scales)
library(janitor)
library(haven)
library(labelled)
library(tidybayes)
library(ggdist)
library(countrycode)
library(lubridate)
library(brms)
library(patchwork)
library(knitr)
library(ggstance)
library(bayestestR)
library(ggh4x)
library(tidyverse)

opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  include = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)
# For tables: Show NAs as blanks
options(knitr.kable.NA = '')

# Create directory for intermediate files
dir.create("data", FALSE)

# Plotting options
theme_set(
  theme_linedraw(
    base_size = if_else(knitr::is_latex_output(), 9, 8)
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.75)),
      axis.ticks = element_line(size = rel(0.3))
    )
)
```

```{r functions}
# Function to replace extreme percentages
percent2 <- function(x, accuracy = .1) {
  x <- percent(x, accuracy = accuracy)
  x <- if_else(x == "100.0%", ">99.9%", x)
  x <- if_else(x == "0.0%", "<0.1%", x)
  x
}
```

@@

# First liberate the data from the slow SPSS file and save to disk
gwp_path <- "data-raw/Gallup/gwp-processed.rds"
if (!file.exists(gwp_path)) {
  gwp <- read_spss(
    "data-raw/Gallup/The_Gallup_101521.sav",
    col_select = c(
      YEAR_CALENDAR,
      COUNTRYNEW,
      WPID, WP1220, WP1219,
      WP16,
      WP60, WP61, WP63, WP65, WP67,
      WP68, WP69, WP70, WP71, WP74
    )
  )
  # Get rid of SPSS attributes
  gwp <- gwp %>%
    zap_labels() %>%
    zap_label() %>%
    zap_widths() %>%
    zap_formats()
  # Save into a good format
  write_rds(gwp, gwp_path)
} else {
  gwp <- read_rds(gwp_path)
}

@@

# Scale scores note rescaling to percentages
gwp <- gwp %>%
  mutate(
    Life_satisfaction = Life_satisfaction * 10,
    Negative_experiences =
      (rowMeans(select(., wp68:wp74), na.rm = TRUE) - 1) * 100,
    Positive_experiences =
      (rowMeans(select(., wp60:wp67), na.rm = TRUE) - 1) * 100
  ) %>%
  select(-c(wp60:wp74))

@@

@@

gwp_mock_path <- "data-raw/gwp-MOCK.rds"
if (dir.exists("data-raw/Gallup") & !file.exists(gwp_mock_path)) {
  library(synthpop)
  gwp_syn <- gwp %>% 
    drop_na()
  gwp_syn <- syn(
    gwp_syn, 
    method = c("", "", "", "", "", "", "normrank", "normrank"),
    maxfaclevels = 168
  )
  gwp_syn <- tibble(gwp_syn$syn)
  saveRDS(gwp_syn, gwp_mock_path)
}

# Load if actual data doesn't exist
if (!dir.exists("data-raw/Gallup")) {
  cat("Using synthetic GWP data.")
  gwp <- read_rds(gwp_mock_path)
} else {
  cat("Using actual GWP data.")
}

# Load data files and merge to one table
gbd <- list.files(
  "data-raw/GBD/",
  pattern = ".zip", full.names = TRUE, recursive = TRUE
) %>%
  read_csv()

# First step is to clean names and remove unnecessary variables
gbd <- clean_names(gbd)
# distinct(gbd, measure, metric)
gbd <- select(gbd, -measure, -metric)

# Rates to percentages
gbd <- gbd %>%
  mutate(across(c(val, upper, lower), ~ .x / 1000))

@@

# Harmonise old names and replace only if harmonised name found
gbd <- gbd %>%
  mutate(
    country_harmonised = countryname(
      country,
      destination = "cldr.short.en"
    )
  ) %>%
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>%
  select(-country_harmonised)

# Download data sets from the ITU website if not yet downloaded
if (!file.exists("data-raw/ITU/PercentIndividualsUsingInternet.xlsx")) {
  dir.create("data-raw/ITU")
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2022/July/PercentIndividualsUsingInternet.xlsx",
    "data-raw/ITU/PercentIndividualsUsingInternet.xlsx"
  )
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2022/July/MobileBroadbandSubscriptions_2007-2021.xlsx",
    "data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx"
  )
}

# Read and reshape mobile broadband data
itu_mobile <- read_xlsx(
  "data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx",
  sheet = 2,
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Economy)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(mobile = as.numeric(value)) %>%
  select(country = Economy, year, mobile)

# Percent using internet
itu_percent <- read_xlsx(
  "data-raw/ITU/PercentIndividualsUsingInternet.xlsx",
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Economy)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(internet = as.numeric(value)) %>%
  select(country = Economy, year, internet)

# Merge to one table in wide format (oldest on left to include all years)
itu <- reduce(list(itu_percent, itu_mobile), left_join)

# years to numbers
itu <- itu %>%
  mutate(year = as.numeric(year))

# Drop tables
rm(itu_mobile, itu_percent)

@@

# Summary
itu_sum <- itu_long %>%
  filter(!(name=="mobile" & year < 2007)) %>% 
  group_by(Metric = name) %>%
  summarise(
    total = n(), 
    total_not_missing = sum(!is.na(value)),
    missing = sum(is.na(value)),
    imputed = sum(is.na(value) & !is.na(value_i))
  )

@@

@@

# Center year
dat <- dat %>%
  mutate(year = (year - 2010))

@@

@@

# Selfharm models don't converge with SEs
dat <- dat %>% 
  mutate(se = if_else(outcome == "Selfharm", 0, se))

# Save data for models
saveRDS(dat, "data/data-all.rds")
```

In 2005, an estimated 17% of the global population used the internet, but by 2020 this number was already 59% [@ituITUICTStatistics2021]. Accompanying the rapid spread of the internet, worries have proliferated that its broad adoption—and technologies enabled by it such as online games, smartphones, and social media—is actively harming its users, particularly adolescents [@carrShallowsHowInternet2010; @turkleAloneTogetherWhy2011]. As a response, users and a growing number of national governments have acted in limiting access to online technologies [@buckleyChinaTightensLimits2021; @departmentfordigitalculturemedia&sportDraftOnlineSafety2021; @kattulaGamePoliciesContexts2021].

However, evidence for widespread harms of online technologies is limited. Initial reports of internet-facilitated harms have been challenged by later work informed by many methodologies including longitudinal [@jensenYoungAdolescentsDigital2019], specification curve analyses [@orbenAssociationAdolescentWellbeing2019], meta-analyses [@appelAreSocialMedia2020], and systematic reviews [@bestOnlineCommunicationSocial2014; @dickson2018screen; @odgersAnnualResearchReview2020; @ophirNewMediaScreenTime2020]. Reviews indicate early research has been hampered by inaccurate measurements of engagement with internet and related technologies [@davidsonFuzzyConstructsTechnology2022; @parrySystematicReviewMetaanalysis2021; @scharkowAccuracySelfReportedInternet2016], biased convenience samples drawn predominantly from countries in the Global North [@ghaiLackSampleDiversity2021], studying a limited range of well-being outcomes [@orbenSocialMediaEnduring2019], and reliance on self-reported evaluations in place of clinical estimates of important mental health outcomes [@campbellInternetUseSocially2006]. A comprehensive test of the overall association between internet adoption and well-being and mental health, broadly defined, has therefore not been conducted. As a consequence, much of the evidence purporting to show that internet technology adoption is associated with broad changes in well-being and mental health, or even causes negative outcomes, remains equivocal [@dickson2018screen; @hawkesCMOReportUnable2019].

In this research we present two studies of global well-being and mental health in the internet age. In the first, we focus on three aspects of psychological well-being, and contrast them with yearly per-capita internet users and mobile broadband subscriptions, across 168 countries and 16 years. In the second we examine three mental health outcomes across 202 countries and 19 years. Our aim is to better understand: 1) How well-being and mental health have changed, on a global scale, during the past two decades of dramatic proliferation of internet technologies and connectivity; 2) How per-capita internet users and mobile broadband subscriptions predict country-level well-being and mental health within a given country and across countries; and 3) Assess the extent to which associations between internet technology adoption and well-being and mental health differ across age and sex and if they are specific to previously suggested vulnerable populations, such as young women.

# Study 1: Well-being

```{r gwp-data-summaries}
gwp_sum <- gwp %>% 
  group_by(outcome) %>% 
  summarise(n = sum(n), cells = n()) %>% 
  mutate(across(n:cells, ~number(., big.mark = ",")))
```

In Study 1, we focused on psychological well-being, as reflected in self-reports of life satisfaction, positive experiences, and negative experiences from `r gwp_sum$n[1]` individuals aged 15 to 89 across `r length(unique(gwp$country))` countries from `r paste0(range(gwp$year), collapse = " to ")`. We contrasted those observations with time series data of the countries' per capita internet users and mobile broadband subscriptions [@ituITUICTStatistics2021] to examine if and how internet and mobile broadband adoption predicted psychological well-being over the past two decades.

## Transparency and Openness

This study was not preregistered. All code used in this project, along with the GBD data, link to the ITU data, and a synthetic version of the GWP data are available at <https://doi.org/10.5281/zenodo.7004053>. Our sample size was determined by the number of available years, countries, and demographics in each respective dataset.

## Methods

### Internet adoption

We identified the International Telecommunication Union's (ITU) database of information and communications technology (ICT) as the most comprehensive country-level source of time-series data on internet adoption [@ituITUICTStatistics2021]. The ITU has collated, from 222 countries' statistical and telecommunications agencies, the yearly percentages of population using the internet from 2000 to `r max(itu$year)`, and yearly per capita broadband subscriptions from 2007 to `r max(itu$year)`. Our analyses used `r number(itu_sum[1, 3, drop = TRUE], big.mark = ",")` yearly internet percentages and `r number(itu_sum[2, 3, drop = TRUE], big.mark = ",")` broadband subscription rates; we imputed `r number(itu_sum[1, 5, drop = TRUE], big.mark = ",")` (`r number(itu_sum[2, 5, drop = TRUE], big.mark = ",")` mobile) intermediate missing values to this dataset using linear interpolation without extrapolation.

### Subjective well-being

We examined subjective well-being indicators from the Gallup World Poll (GWP), a nationally representative annual survey of 1,000 civilian, non-institutionalised individuals aged 15 years or older from `r length(unique(gwp$country))` countries from 2005 to 2021. The surveys are conducted face-to-face or via telephone, in the respondents' native language and by local interviewers. For details on the GWP sampling and survey methodology, see [@gallupGallupWorldPoll2020; @gallupHowDoesGallup2014].

GWP measures subjective well-being with the positive and negative experience indices, which measure respondents' experienced well-being on the day before the survey with five items each. For positive experiences, these items are "Did you feel well-rested yesterday?"; "Were you treated with respect all day yesterday?"; "Did you smile or laugh a lot yesterday?"; "Did you learn or do something interesting yesterday?"; and "(Did you experience the following feelings during a lot of the day yesterday?) How about enjoyment?" And for negative experiences, the items are responses to "Did you experience the following feelings during a lot of the day yesterday?" with prompts "How about physical pain?"; "How about worry?"; "How about sadness?"; "How about stress?"; and "How about anger?" We aggregated both scales for each respondent by taking a mean of the five items.

Life satisfaction in the moment was measured with one 11-step Likert item, "Please imagine a ladder, with steps numbered from 0 at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life for you. On which step of the ladder would you say you personally feel you stand at this time?", similar to the Cantril self-anchoring scale [@cantril1965pattern; @kapteynDimensionsSubjectiveWellbeing2015]. For analyses, we converted these variables to percentages, and aggregated the `r gwp_sum[1, 2]` individuals' data to means and standard errors for each outcome, country, year, sex, and age (5-year age groups from 15 to 89 years) combination (`r gwp_sum[1, 3]` cells).

### Data analysis

We analysed the data with meta-analytic bayesian hierarchical regression models [@burknerBrmsPackageBayesian2017; @gelmanDataAnalysisUsing2007; @standevelopmentteamStanModelingLanguage2021]. We first asked how the well-being outcomes had changed over time. To answer, we modelled each outcome (e.g. life satisfaction) on population level intercepts (we centred time on 2010), contrasts of time (in continuous years), sex (Female: 1, Male: -1), and their interaction. We allowed all coefficients to vary across countries, age groups, and the age by country interactions, and treated all outcomes as normally distributed.
